---
title: 'Exploratory Data Analysis Case Study Project #1'
author: "Nitin Seelam, Edmund Tran, Jian Kuang"
date: "December 4, 2015"
output: html_document
runtime: shiny
---

dictionary_URL : "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf"

```{r load_data}
df_data <- read.csv("data/ss14hca.csv")
```

### Nitin
```{r}
hist(df_data$FINCP, main="Family Income", col="Red", xlab="Income")
```

The histogram above shows the distribution of income within California. However the data might be restricted to a maximum amount, for privacy concerns.


```{r,}
boxplot(VALP~BDSP, df_data, col="green", main="Property Value by Number of Bedrooms",xlab="Number of Bedrooms", ylab="Property Value")
```

There is a clear effect on property value by the varying number of bedrooms. It is expected that the more rooms there are in a home, the more valuable the property would be and that is show in the boxplot above. The dark lines in the boxes display the mean value and the box itself displays the 25th to 75th price range, of all properties with a certain number of bedrooms.


```{r,echo=FALSE}
dataMatrix <- na.omit(df_data[,c("VALP", "BDSP", "YBL")])
dataMatrix1<- dataMatrix[1:10000,]
pairs(~BDSP + VALP + YBL, data=dataMatrix1, main="Scatterplot on Housing Factors", col="blue")
```

A sample of complex plots which can be made to look at correlations between multiple predictor variables.

### Edmund

We'll show how effective a sample mean is at estimating the population mean

Sample: monthly rents of x properties
Population: monthly rents of all properties in California (52,358 data points).
Population Mean: $1,249


```{r, echo=FALSE}
inputPanel(
  sliderInput("sample_size", label = "Sample Size:",
              min = 10, max = 100, value = 10, step = 10)
)

renderPlot({
  meanV <- NULL
  allRents <- df_data[complete.cases(df_data$RNTP), "RNTP"]
  # choose sample 1000x
  for (i in 1:1000) {
    sampleRents <- sample(allRents, input$sample_size)
    meanV <- c(meanV, mean(sampleRents))
  }

  hist(meanV,
     xlab = "Sample Mean",
     ylab = "Frequency",
     xlim = c(600, 1800),
     ylim = c(0, 400),
     main = "Distribution of 1000 Sample Means of Monthly Rents in CA")
  
  # show mean on graph for reference
  abline(v = mean(allRents), col = "red")
})
```


## Cluster : Jack Kuang
### 2 numeric variables cluster : property value(VALP) and insurance(INSP)  
let's see the summary of these two variables first:
```{r, echo=FALSE}
par(mfrow=c(1,2))
boxplot(df_data$VALP)
boxplot(df_data$INSP)
```

We use shiny package at here, so customer can change the number of cluster
```{r cluster_2vars, echo=FALSE}
inputPanel(
  selectInput("n_clusters", label = "Number of clusters:",
              choices = c(2, 3, 5, 10), selected = 3)
)

## cluster property value(VALP), insurance(INSP)
dataMatrix <- na.omit(df_data[,c("VALP", "INSP")])
dataMatrix.scaled <- scale(dataMatrix)

renderPlot({
    kmeansObj <- kmeans(dataMatrix.scaled, centers = as.numeric(input$n_clusters))
    plot(dataMatrix, col=kmeansObj$cluster, main = "cluster on property value(VALP) and insurance(INSP)", 
         xlab="property value(VALP)", ylab = "insurance(INSP)")
})
```

### 3 numberic variables cluster : property value(VALP), insurance(INSP), year built(YBL)
```{r cluster_3vars, echo=FALSE}
## cluster property value(VALP), insurance(INSP), year built (YBL)
dataMatrix2 <- na.omit(df_data[,c("VALP", "INSP", "YBL")])
dataMatrix2.scaled <- scale(dataMatrix2)
kmeansObj2 <- kmeans(dataMatrix2.scaled, centers = 3)
plot(dataMatrix2, col=kmeansObj2$cluster)
```
