---
title: 'Statistical Modeling and Simulation'
author: "Nitin Seelam, Edmund Tran, Jian Kuang"
date: "December 10, 2015"
output: html_document
runtime: shiny
---

## Loading Data
dictionary_URL : "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf"

Because the csv file's size makes it inconvenient to load, we cache it into a single R object (rds file).
```{r load_data, cache=TRUE}
# df_data <- read.csv("data/ss14hca.csv")
# saveRDS(df_data, file = "data/ss14hca.rds")
df_data <- readRDS("data/ss14hca.rds")
```

## Basic Explortary Plots
### California Family Income Histgram
```{r}
hist(df_data$FINCP, main = "Family Income", col = "Red", xlab = "Income")
```

The histogram above shows the distribution of income within California. However the data might be restricted to a maximum amount, for privacy concerns.

### How does property value change with the number of bedrooms?
```{r,}
boxplot(VALP~BDSP, df_data, col = "green", main = "Property Value by Number of Bedrooms", xlab = "Number of Bedrooms", ylab = "Property Value")
```

There is a clear effect on property value by the varying number of bedrooms. It is expected that the more rooms there are in a home, the more valuable it will be and that is shown in the boxplot above. The dark lines in the boxes display the mean value and the box itself represents the 25th to 75th percentile of property values for all properties with a certain number of bedrooms.

### Relationship between Property Value and Family Income 
We add a linear regression line to the plot to the relationship.
```{r}
with(df_data, smoothScatter(VALP, FINCP, xlab = "Property Value", ylab = "Family Income"))
model <- lm(FINCP ~ VALP, df_data)
abline(model, lwd=2, col="green")
```

### Correlation Metrics : Property Specifications 
BDSP : bedroom number  
VALP : property value  
YBL : year built  
```{r,echo=FALSE}
dataMatrix <- na.omit(df_data[,c("VALP", "BDSP", "YBL")])
dataMatrix1<- dataMatrix[1:10000,]
pairs(~BDSP + VALP + YBL, data=dataMatrix1, main="Scatterplot on Housing Factors", col="blue")
```

This is a sample of the complex plots that can be made to look at correlations between multiple predictor variables.

## Statistical Inference and Simulation

We'll show how effective a sample mean is at estimating the population mean.

Sample: monthly rents of x properties
Population: monthly rents of all properties in California (52,358 data points).
Population Mean: $1,249

Move the sample size slider to see how bigger sample sizes give a better estimate of the population mean.

```{r, echo=FALSE}
inputPanel(
  sliderInput("sample_size", label = "Sample Size:",
              min = 10, max = 100, value = 10, step = 10)
)

renderPlot({
  meanV <- NULL
  allRents <- df_data[complete.cases(df_data$RNTP), "RNTP"]
  # choose sample 1000x
  for (i in 1:1000) {
    sampleRents <- sample(allRents, input$sample_size)
    meanV <- c(meanV, mean(sampleRents))
  }

  hist(meanV,
     xlab = "Sample Mean",
     ylab = "Frequency",
     xlim = c(600, 1800),
     ylim = c(0, 400),
     main = "Distribution of 1000 Sample Means of Monthly Rents in CA")
  
  # show mean on graph for reference
  abline(v = mean(allRents), col = "red")
})
```


## Cluster
###Clustering on two numeric variables: property value and insurance price  
Let's see the summary of these two variables first:
```{r, echo=FALSE}
par(mfrow=c(1,2))
boxplot(df_data$VALP, main="Property Value")
boxplot(df_data$INSP, main="Insurance Price")
```

We use the shiny package here, so you can change the number of clusters.
```{r cluster_2vars, echo=FALSE}
inputPanel(
  selectInput("n_clusters", label = "Number of clusters:",
              choices = c(2, 3, 5, 10), selected = 5)
)

## cluster property value(VALP), insurance(INSP)
dataMatrix <- na.omit(df_data[,c("VALP", "INSP")])
dataMatrix.scaled <- scale(dataMatrix)

renderPlot({
    kmeansObj <- kmeans(dataMatrix.scaled, centers = as.numeric(input$n_clusters))
    plot(dataMatrix, col=kmeansObj$cluster, main = "Cluster on Property Value and Insurance Price", 
         xlab="Property Value", ylab = "Insurance Price")
})
```

###Clustering on 3 dimensions: property value (VALP), insurance (INSP), year built (YBL)
```{r cluster_3vars, echo=FALSE, cache=TRUE}
## cluster property value(VALP), insurance(INSP), year built (YBL)
dataMatrix2 <- na.omit(df_data[,c("VALP", "INSP", "YBL")])
dataMatrix2.scaled <- scale(dataMatrix2)
kmeansObj2 <- kmeans(dataMatrix2.scaled, centers = 3)
plot(dataMatrix2, col=kmeansObj2$cluster)
```

